
```{r, echo = FALSE, results = 'hide', message = FALSE}
library(tidyverse)
COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")
```


```{r, echo = FALSE, results = 'hide', message = FALSE}
justthecarrier <- COflights %>%
  filter(DEST == "DEN") %>%
   select(CARRIER) %>%
  distinct()
```




**Probability each carrier is delayed given all of its flights to Denver**
```{r, echo = FALSE, message = FALSE}
probabilityplot <- COflights %>%
  filter(ARR_DELAY >= 15, DEST == "DEN") %>%
  group_by(CARRIER) %>%
  summarize(count = n()) %>%
  select(count)
totalpercarrier <- COflights %>%
  filter(DEST == "DEN") %>%
  group_by(CARRIER) %>%
  summarize(count = n()) %>%
  select(count)
ProbabilityOfDelay <- probabilityplot/totalpercarrier
```

**A plot showing which carrier has the highest probability of delay given the amount of flights they had into denver.**

```{r, echo = FALSE, message = FALSE}
tableplot <- cbind(justthecarrier, ProbabilityOfDelay) %>%
  distinct()
tableplot


ggplot(data = tableplot) +
  geom_col(mapping = aes(x = CARRIER, y = count, fill = CARRIER)) +
  ylab('Probability of delay')
```





**Calculating the joint probability of plane arriving late(to denver) and that plane being 1,000 miles or greater a distance**

```{r, echo = FALSE, message = FALSE}
jointprobability <- COflights %>%
  filter(ARR_DELAY >= 15, DISTANCE >= 1000, DEST == "DEN") %>%
  summarise(count = n())

TotalJoinProbability <- COflights %>%
  filter(ARR_DELAY >= 15 | DEST == "DEN") %>%
  summarise(count = n())

FinalProbability <- jointprobability/TotalJoinProbability
FinalProbability

```



**probability a flight is late**
```{r, echo = FALSE, message = FALSE}
Lateflight <- COflights %>%
  filter(ARR_DELAY >= 15) %>%
  summarize(count = n())

Allflights <- COflights %>%
  count(CANCELLED != 1) 
y <- 473153

probabilityitslate <- Lateflight/473153
probabilityitslate


```







**Vague Question and a table to answer**
```{r, echo = FALSE, message = FALSE}
Weatherdelays <- COflights %>%
  filter(DEST == "DEN") %>%
  select(WEATHER_DELAY) %>%
  distinct() %>%
  filter(WEATHER_DELAY >= 15, WEATHER_DELAY != 'NA') %>%
  summarise(count = n())
NASdelays <- COflights %>%
  filter(DEST == "DEN") %>%
  select(NAS_DELAY) %>%
  distinct() %>%
  filter(NAS_DELAY >= 15,NAS_DELAY != 'NA') %>%
  summarise(count = n())
LateAirDelay <- COflights %>%
  filter(DEST == "DEN") %>%
  select(LATE_AIRCRAFT_DELAY) %>%
  distinct() %>%
  filter(LATE_AIRCRAFT_DELAY >= 15, LATE_AIRCRAFT_DELAY != 'NA') %>%
  summarise(count = n())
SecurityDelay <- COflights %>%
  filter(DEST == "DEN") %>%
  select(SECURITY_DELAY) %>%
  distinct() %>%
  filter(SECURITY_DELAY >= 15, SECURITY_DELAY != 'NA') %>%
  summarise(count = n())
CarrierDelay <- COflights %>%
  filter(DEST == "DEN") %>%
  select(CARRIER_DELAY) %>%
  distinct() %>%
  filter(CARRIER_DELAY >= 15, CARRIER_DELAY != 'NA') %>%
  summarise(count = n())


Delayplot <- rbind(Weatherdelays,CarrierDelay,SecurityDelay,LateAirDelay,NASdelays)
m <- matrix(1:1, nrow = 5)
rownames(m) <- c('Weather','Carrier','Security','Lateaircraft','NAS')
finaltable <- cbind(m,Delayplot)
finaltable[1] <- NULL
finaltable
```



**Answer to my individual**
Two factors I discovered that affect the arrival in denver are the carrier, which can increase the delay probability, or the distance which can also increase the probability of a delay, although it is not a large increase.

**Answer to Vague Question**
From the table I created above, we can clearly see that Carrier delays have the greatest affect on late flights, while security bare almost no effect.
**What I Did**
For this lab, I completed the full individual portion.  Furthermore, I created a table that calculated the total number for each type of delay, when it resulted in a flight being late to denver. I then created a 1:5 Matrix with the delay name types, and binded this column to the table of data.  Using this, I was  able to answer the vague question about what causes the most late flights.  I did alot of data manipulation in order to create my plot as well.