```{r}
library(tidyverse)
COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")
COflights
```

Probability of AA that were cancelled

```{r}
indiarr <- COflights%>%
  filter(!is.na(ARR_TIME))
  filter(ORIGIN=="DEN"|DEST=="DEN")
indiarr
```


```{r}
filter(indiarr, CARRIER=="AA")%>%
  summarise(indiarr = n())
filter(indiarr, CARRIER=="AA"&CANCELLED ==1)%>%
  summarise(indiarr = n())
filter(indiarr, CARRIER=="AA"&CANCELLED ==1)%>%
  summarise(indiarr = n())/filter(indiarr, CARRIER=="AA")%>%
  summarise(indiarr = n())
```
What is the probability of flight on 2017/8 and origined at denver?

```{r}
filter(indiarr, YEAR==2017&MONTH ==8)%>%
  summarise(indiarr = n())
filter(indiarr,YEAR==2017&MONTH ==8&ORIGIN=="DEN" )%>%
  summarise(indiarr = n())
filter(indiarr,YEAR==2017&MONTH ==8&ORIGIN=="DEN" )%>%
  summarise(indiarr = n())/filter(indiarr, YEAR==2017&MONTH ==8)%>%
  summarise(indiarr = n())

```

What is the probability that the flight is delaid given that the carrier is DL?

```{r}
filter(indiarr, DEP_DELAY>=0)%>%
  summarise(indiarr = n())/nrow(indiarr)
filter(indiarr, CARRIER == "DL")%>%
  summarise(indiarr = n())/nrow(indiarr)
filter(indiarr, DEP_DELAY>=0)%>%
  summarise(indiarr = n())/nrow(indiarr)*filter(indiarr, CARRIER == "DL")%>%
  summarise(indiarr = n())/nrow(indiarr)

```


Chart that shows the probability of delaid flight over the year.


```{r}
chartfile <-filter(indiarr, DEP_DELAY>=0)%>%
  group_by(MONTH)%>%
  summarise( n())/
group_by(indiarr, MONTH)%>%
  summarise(n())
```

Visualization for my finding:


```{r}
monthdelay <- indiarr %>%
  filter(DEP_DELAY >= 0) %>%
  group_by(MONTH) %>%
  summarise(delay = n())

totaldelay <- indiarr %>%
  group_by(MONTH) %>%
  summarise(totaldelay = n()) 

monthdelay$delaycount <- totaldelay$totaldelay

delayprob <- monthdelay %>%
  mutate(prob = delay/delaycount)

```


```{r}
ggplot(data = delayprob, aes(x = MONTH, y = prob))+
  geom_point()+
  geom_smooth()
```




Findings:1, The carrier "AA" never got canceled in the dataset.
2, flights on 2017/8 were mostly originated from Denver(47%)
3, Carrier "DL" had a low chances to be deliad, should be rewarded!!!!.
4. January's flights had the highest percentage being delaid. While November had lowest percentage being delaid.


What I have done: Helped my team to figure out the way to calculate the probability, did my individual part, created a visualization for my team, generated a research question, found the answer for that question.



Team part:

Which Carrier delaid the most often ?

```{r}
Team <-filter(indiarr, DEP_DELAY>=0)%>%
  group_by(MONTH, CARRIER)%>%
  summarise(total = n())

```

```{r}
ggplot(data = Team) + 
  geom_point(mapping = aes(x = MONTH, y = total)) + 
  facet_wrap(~ CARRIER, nrow =5)
```
 Finding: Obviously, WN has the highest rate to be delaid comparing to other airlines, and AS, B6, and EV were very unlikely to be late over the time.
 
 Why knowing this is important? Because the airline should know how good or terrible their services are, and knowing this may help them earn their customers back. 
